Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-09-06 Paper.qmd
Latest edits.

Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996]. In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971]. At its simplest, this would have taken the form of mothers providing investment to their children.

The fitness interdependence of offspring and at least one parent is seen in many animals [@clutton-brock1991; @webb1999]. However, humans are a uniquely interdependent species among the living primates [@tomasello2012] in that one’s fitness prospects depend greatly on a network of kin and other social allies [@apicella2019]. Although also true with adults, this need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015]. Much of this investment comes from parents [@kramer2018], with most human reproduction occurring in the context of biparental investment and long-term pair bonds [@apostolou2007; @schacht2019]. However, in many societies, other adults like grandparents, aunts, uncles, and non-relatives have also been found to provide significant investment in children [@kramer2018; @sear2008], something which was likely favored due to indirect fitness benefits, reciprocity, or reputation management [@gurven2000] + (Schniter Preprint (permission needed). Children also often serve as alloparents, helping to raise their younger siblings or provide for them through food preparation or collection [@hrdy2009; @meehan2009; @sear2008], despite still requiring substantial amounts of investment themselves [@kramer2018a].

@@

Although the fitness prospects of both children and familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974]. Most resources that translate to fitness benefits are limited so that one unit of a resource invested in one area cannot be used in another. In response to this basic fact of life, humans, like other organisms, are expected to have evolved to efficiently allocate resources between somatic effort (i.e., growth and maintenance) and reproductive effort (i.e., mating and parenting effort) [@daly1983], something which often pits the fitness interests of parents against the fitness interests of offspring.

@@

Conflicts of interest between children and parents may also involve how much children invest in the household. As previously mentioned, each individual child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974]. This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food [@lowe2003; @syme2023a]. Conflicts of interest over investment in the household also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more care for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household if the child’s presence limits how much caretakers can invest in other children they deem more worthy of investment [@kushnick2023].

@@

When conflicts of interest over investment exist, there may be much to gain from deception (i.e., exaggerating or completely fabricating one’s quality or need) so that one can receive the benefits that come from being seen as having a particular quality without actually doing so [@zahavi1993]. Although the risks of being deceived can be substantial, signal receivers may still benefit from paying attention to signals so long as the information is credible enough to inform one’s behavior in ways that are adaptive on average [@bliegebird2001; @grafen1990]. These benefits coupled with the likelihood of strong selection against individuals who respond to fake signals results in the expectation that signals within species will often need to benefit both signaler and receiver to persist evolutionarily [@hasson1994; @maynard-smith2004].

@@

Regardless of the mechanism behind the credibility of a signal, credible signals work by revealing private information about one’s qualities to others. In the context of child signals of need, a child has access to information about their internal state, which is unavailable to others, that allows them to estimate their valuation of additional investment. By signaling in a credible way, the child reveals their resource valuation to potential caretakers who may be skeptical of the child’s claims due to the common conflicts of interest described above. With the new information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing. When this is the case, all parties are better off than they would be without the signal (so long as the benefits outweigh the costs of the signal to child and parent), with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response.

@@

Potentially combine and condense with paragraph below: As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004). For this reason, there is likely a transition period in which crying is primarily meant to provide information about how much a child values further investment. Many models have been put forward which attempt to explain human crying and begging in other species as credible signals of need due to the differential costs and benefits of crying and analogous vocalizations in other species as deterring the exaggeration of need [@godfray1991; @kilner1997; @parker2002]. Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].

@@

The survey data for this study come from a larger project run by the Utila Child Health Project. Data was collected by Jess Hlay, Izabel Rodriguez, Daniela Mejia, Angela Garcia, or Caroline Smith who read the survey questions to participants and recorded their answers. Participants could choose to do the survey in English or Spanish.

@@

Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness. We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create the CurrentHealthMean variable. In the `r sum((!is.na(modeldf$Currenthealth2)) & is.na(modeldf$Currenthealth3))` instance in which we only data for 2 of the measures, we still calculated the mean a numerator of 2. See Table @tbl-healthvars for the wording of these variables.

@@

The survey also included many demographic questions about the survey taker and their households. Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide allocare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income. This data allowed us to calculate the number of children (total, younger, and older) and adults within the household.

@@

To model the effects of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011]. This resulted in variables which represented how each child differed from the average child for their given sex and age.

@@

All statistical analyses were exploratory and involved fitting multiple models while being cognizant of the risk of overfitting the data, but with the payoff of idenftifying theoretically relavent effecrs that should be tested in the future.

We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables. As our signaling variables invovled over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with household as a random effect. Control variables differed in some models, but always include child age and sex as controls. Specific control variables are mentioned in the results

@@

`r dplyr::n_distinct(modeldf$householdID)` primary caretakers took the survey and provided signaling data for at least one child. `r nrow(modeldf)` children had frequency data for at least 1 signal, with `r nrow(modeldf_FULLSIG)` having data for all 3. Among those with at least one record of signaling frequency `r sex_table[1]` were female and `r sex_table[2]` were male. Anthropometric and signaling data exist for `r nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `r sex_table2[1]` were female and `r sex_table2[2]` were male. For summary statistics of child-level variables see @tbl-sumtab_children. For summary statistics of caretaker and household-level variables see @tbl-sumtab_adults.

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child study variables.
vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures",
  HeightZ = "Height Z-score (CDC)",
  WeightZ = "Weight Z-score (CDC)",
  BMIZ = "BMI Z-score (CDC)",
  SubscapMean = "Subscapular skinfold thickness",
  SubscapR = "Subscapular skinfold residuals",
  TricepMean = "Tricep skinfold thickness",
  TricepR = "Tricep skinfold residuals",
  BodyFat =  "Body fat composite"
)

@@

Alluvial plot showing the combinations of signaling frequencies per month. Alluvium width equals the number of individiuals with the same frequencies for the three signals/states. Stratum height equals the count of individuals at a given frequency. Frequency key: Low includes "never" and "once a month or less," mediums includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and "multiple times per day."
:::

@@

The effect of age on specific signal frequencies (per month), summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-main_models for the corresponding regression table.
:::

@@

The effect of the frequency of conflict with one's primary caretaker (per month) on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-main_models for the corresponding regression table.
:::

@@

The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). See @tbl-main_models for the corresponding regression table.
:::

@@

The effect of A) child age and B) the number of adults in the household that contribute to childcare on the frequency of conflict between caretaker and child. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-conflict_dep_var for the corresponding regression table.
:::

@@

The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Rug corresponds to actual values of child alloparenting frequency with a small amount of jitter to reveal overlapped values. See @tbl-conflict_dep_var for the corresponding regression table.
:::

@@

## Positive assoication between child alloparenting effort and number of older children

@@

::: {#fig-effect_bodyfat}
```{r}
signaling_data_plot_bodyfat
```

@@

Plots showing the likelihood of child being more needy, as needy, or less needy than a caretaker's other children. A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
:::

@@

Plots showing the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried. A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.) Legend key: -1 = negative response, 0 = neutral response, 1 = positive response (all from child's perspective).
:::

@@

As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, something which is likely due to numerous reasons. One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of older children and adults. Another is that as children survive to the point where they would have been unlikely to die young in ancestral environments, the intensity conflicts of interest with parents decrease due to children passing this important milestone and requiring less investment to reach adulthood than younger children, current or future (cite daly and wilson). For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\]. The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned. It is also possible that parents see older children less than younger children. However, the negative assocition between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.

There was no significant association between the frequency of child sadness and age. Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects (cite), from which, others can still obtain cue-based information about one's emotional state, well-being, and desires. By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals. For this reason, a change in signals that arise from sadness outside of crying (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often. It is also possible that there is no effect of age on sadness frequency among the age ranges examined it.

-   add in depression over the lifecourse

@@

We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency of and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys. In many societies, girls are often expected to do more alloparenting within the household and often end up doing so (cite). While girls in this sample were also reported to give more allocare, the difference was not significant. However, Utilian mothers reported in focus groups (confirm) that girls tend to provide more help for their siblings than boys. If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.

@@

Children with less body fat cried more, threw more temper tantrums, and had higher summed signal costs. In other words, needier children signaled more often. Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys. Female reproduction is highly dependent on body fat, with conception requiring about xx% body fat and mothers investing a significant amount of calories in both children in the womb and breastfed infants (cite). For these reasons, low body fat is likely a be better predictor of future reproductive success in older girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.

@@

Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not. One potential causal explanation for this is that parents may often have information for sources outside of child signaling to corroborate a child's need (e.g., a fever, a diagnosis, or witness a child hurt themselves) so that there is less private information about a child's current state. Simultaneously, conflicts of interest may be less significant in these times due to either 1) both sides benefiting similarly from their child returning to normal (e.g., with an illness) or 2) the cause involving something which is minor and would quickly resolve itself (e.g. a skinned knee). It may also be that many of the typical positive responses to pain and illness are not very costly (e.g., providing comfort).

When the cause involved punishment, however, parents were more likely to respond negatively and less likely to respond positively. In these circumstances, children may be signaling about the severity of a punishment (cite). However, parents may be reluctant to reduce punishments if the behavior that warranted the punsishment was repeated or severe. They may also react with scolding or further punishment. That said, our model predicts positive responses in xxxxx cases, suggesting that children are successful in negotiating better outcomes at least some of the time. - costly apolgy model?

@@

*CUT Negative association between caretaker estimates of child health and child signaling for some measures. This does not extend to the costlier signals.*

@@

Monthly income was also not found to be a significant predictor of signaling behavior. Many models designed to explain signals of need (e.g., the begging of young birds or withdrawal in humans) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more (cite). However, it is possible that this can be true and household level resources have little influence on child signaling frequencies between families. As signaling behavior is costly to child, caretakers, and siblings, children may gain direct and indirect fitness benefits from limiting their signaling behavior in times of resource stress (i.e., when the chances of gaining investment or the size of investment are small). Similarly, children in high resource families may also be motivated to signal at higher than expected frequencies due to the potential for larger pieces of investment. We know of no empirical explorations of this in humans (check more and confirm). However, multiple game theoretical models of sibling competition and have found that parental resources have a small impact on offspring signaling, especially with high rates of monogamy \[Bossan + Mock and Parker + Check for newer sources\]. Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.

@@

Another limitation is that our study was cross-sectional. A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average. In other words, higher levels of signaling may stem from greater levels of need while simultaneously eliciting more investment from others. This hypothesized causal, bi-directional relationship makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age). Longitudinal studies can help address this. However, our small sample size of participants with signaling data and anthropometric data at two time periods prevents this for this study.

@@

>>>>

Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996]. In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971].

Care motivated by a high degree of fitness interdependence between offspring and at least one parent is seen in many animals [@clutton-brock1991; @webb1999]. However, humans are a uniquely interdependent species among the living primates [@tomasello2012] in that one’s fitness prospects depend greatly on a network of kin and other social allies [@apicella2019]. This need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015]. Much of this investment comes from parents [@kramer2018], However, other adults like grandparents, aunts, uncles, and non-relatives have also been found to provide significant investment in children in many societies [@kramer2018; @sear2008], something which was likely favored due to a combination of indirect fitness benefits, reciprocity, or reputation management [@gurven2000] + (Schniter Preprint (permission needed). Children also often serve as alloparents, helping to raise their younger siblings or provide for them through food preparation or collection [@hrdy2009; @meehan2009; @sear2008], despite still requiring substantial amounts of investment themselves [@kramer2018a].

@@

Although the fitness prospects of both children and familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974]. Most resources relevant to fitness outcomes are limited so that one unit of a resource invested in one area cannot be used in another. In response to this basic fact of life, humans, like other organisms, are expected to have evolved to efficiently allocate resources between somatic effort (i.e., growth and maintenance) and reproductive effort (i.e., mating and parenting effort) [@daly1983], something which often pits the fitness interests of parents against the fitness interests of offspring.

@@

Conflicts of interest between children and parents may also involve how much children invest in the household. As previously mentioned, each child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974]. This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food [@lowe2003; @syme2023a]. Conflicts of interest over investment in the household also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more care for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household if the child’s presence limits how much caretakers can invest in other children they deem more worthy of investment [@kushnick2023].

@@

When conflicts of interest over investment exist, there may be much to gain from deception (i.e., exaggerating or completely fabricating one’s quality or need) so that one can receive the benefits that come from being seen as having a particular quality without actually doing so [@zahavi1993]. Although the risks of being deceived can be substantial, signal receivers may still benefit from paying attention to signals so long as the information is credible enough to inform one’s behavior in ways that are adaptive on average [@bliegebird2001; @grafen1990]. These benefits coupled with the likelihood of strong selection against individuals who respond to fake signals results in the expectation that signals within species will often need to benefit both signaler and receiver to persist evolutionary [@hasson1994; @maynard-smith2004].

@@

Credible signals work by revealing private information about one’s qualities to others. In the context of child signals of need, a child has access to information about their internal state, which is unavailable to others, that allows them to estimate their valuation of additional investment. By signaling in a credible way, the child reveals their resource valuation to potential caretakers who may be skeptical of the child’s claims due to the common conflicts of interest described above. With the information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing. When this is the case, all parties are better off than they would be without the signal, with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response as long as the benefits outweigh the costs of the signal to child and parent.

@@

Potentially combine and condense with paragraph below: As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004 comment). For this reason, there is likely a transition period in which crying is primarily meant to provide information about how much a child values further investment. Many models have been put forward which attempt to explain human crying and begging in other species as credible signals of need due to the differential costs and benefits of crying and analogous vocalizations in other species as deterring the exaggeration of need [@godfray1991; @kilner1997; @parker2002]. Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].

@@

The survey data for this study come from a larger project run by the Utila Child Health Project. Data was collected by either Jess Hlay, Izabel Rodriguez, Daniela Mejia, Angela Garcia, or Caroline Smith who read the survey questions to participants and recorded their answers. Participants could choose to do the survey in English or Spanish.

@@

Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness. We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create the CurrentHealthMean variable. In the `r sum((!is.na(modeldf$Currenthealth2)) & is.na(modeldf$Currenthealth3))` instance in which we only data for 2 of the measures, we still calculated the mean but used a numerator of 2. See Table @tbl-healthvars for the wording of these variables.

@@

The survey also included many demographic questions about the survey taker and their households. Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide childcare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income. This data allowed us to calculate the number of children (total, younger, and older) and adults within the household.

@@

For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011]. The resulting variables which represent how each child differed from the average child for their given sex and age.

Summed body fat

@@

All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables while being cognizant of the risk of overfitting the data. The payoff to this is the identification of theoretically relevant effects that should be tested in the future.

We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables. As our signaling variables involved over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with household as a random effect. Control variables differed in some models, but always include child age and sex as controls. Specific control variables are mentioned in the results.

@@

`r dplyr::n_distinct(modeldf$householdID)` primary caretakers took the survey and provided signaling data for at least one child. `r nrow(modeldf)` children had frequency data for at least 1 signal, with `r nrow(modeldf_FULLSIG)` having data for all 3. Among those with at least one record of signaling frequency, `r sex_table[1]` were female and `r sex_table[2]` were male. Anthropometric and signaling data exist for `r nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `r sex_table2[1]` were female and `r sex_table2[2]` were male. For summary statistics of child-level variables see @tbl-sumtab_children. For summary statistics of caretaker and household-level variables see @tbl-sumtab_adults.

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child study variables.
vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures",
  HeightZ = "Height Z-score (CDC)",
  WeightZ = "Weight Z-score (CDC)",
  BMIZ = "BMI Z-score (CDC)",
  SubscapMean = "Subscapular skinfold thickness (mm)",
  SubscapR = "Subscapular skinfold residuals",
  TricepMean = "Tricep skinfold thickness (mm)",
  TricepR = "Tricep skinfold residuals",
  BodyFat =  "Body fat composite"
)

@@

Alluvial plot showing the combinations of signaling frequencies per month. Alluvium width equals the number of individiuals with the same frequencies for the three signals/states. Stratum height equals the count of individuals at a given frequency. Frequency key: Low includes "never" and "once a month or less," medium includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and "multiple times per day."
:::

@@

The effect of age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Rug corresponds to actual values of child age with a small amount of jitter to reveal overlapped values. See @tbl-main_models for the corresponding regression table.
:::

@@

The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Rug corresponds to actual values of the frequency of conflict between child and primary caretaker with a small amount of jitter to reveal overlapped values. See @tbl-main_models for the corresponding regression table.
:::

@@

The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Rug corresponds to actual values of primary caretaker education (years) with a small amount of jitter to reveal overlapped values. See @tbl-main_models for the corresponding regression table.
:::

@@

The effect of A) child age and B) the number of adults in the household that contribute to childcare on the frequency of conflict between caretaker and child. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Rug corresponds to actual values of A) child age and B) the number of adults in the household that contribute to childcare with a small amount of jitter to reveal overlapped values. See @tbl-conflict_dep_var for the corresponding regression table.
:::

@@

The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Rug corresponds to actual values of child alloparenting frequency with a small amount of jitter to reveal overlapped values. See @tbl-conflict_dep_var for the corresponding regression table.
:::

@@

## Positive association between child alloparenting effort and number of older children

@@

::: {#fig-effect_bodyfat}
```{r}
signaling_data_plot_bodyfat2
```

@@

The likelihood of child being more needy, as needy, or less needy than a caretaker's other children. A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
:::

@@

The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried. A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.) Legend key: -1 = negative response, 0 = neutral response, 1 = positive response (all from child's perspective).
:::

@@

As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, something which is likely due to numerous reasons. One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults. Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995]. For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\]. That said, all else may often not be equal if investment in older children is costlier, something which may be reflected in caretakers seeing older children as relatively needy. The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned. ~~It is also possible that parents see older children less than younger children. However, the negative association between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.~~

There was no significant association between the frequency of child sadness and age. Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011], from which, others can still obtain cue-based information about one's emotional state, well-being, and desires. By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals. For this reason, a change in signals that arise from sadness outside of crying (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often. It is also possible that there is no effect of age on sadness frequency among the age ranges examined it.

-   add in depression over the lifecourse?

@@

We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency of and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys. In many societies, girls are often expected to do more alloparenting within the household and often end up doing so [@helfrecht2016 @weisner1977; @barry1971; @turke1988]. While girls in this sample were also reported to give more allocare, the difference was not significant. However, Utilian mothers reported in focus groups (confirm) that girls tend to provide more help for their siblings than boys. If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.

@@

Children with less body fat cried more, threw more temper tantrums, and had higher summed signal costs. In other words, needier children signaled more often. Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys. Female reproduction is highly dependent on body fat, with conception requiring about xx% body fat and mothers investing a significant amount of calories in both children in the womb and breastfed infants [@jasienska2009]. For these reasons, low body fat is likely a be better predictor of future reproductive success in adolescent (and older) girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.

@@

Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not. One potential causal explanation for this is that parents may often have information for sources outside of child signaling to corroborate a child's need (e.g., a fever, a diagnosis, or witness a child hurt themselves) so that there is less private information about a child's current state. Simultaneously, conflicts of interest may be less significant in these times due to either 1) both sides benefiting similarly from their child returning to normal (e.g., with an illness) or 2) the cause involving something which is minor and would quickly resolve itself (e.g. a skinned knee). It may also be that many of the typical positive responses to pain and illness are not very costly to caretakers (e.g., providing comfort), thereby increasing their frequency.

When the cause involved punishment, however, parents were more likely to respond negatively and less likely to respond positively. In these circumstances, children may be signaling about the severity of a punishment (cite syme costly apology). However, parents may be reluctant to reduce punishments if the behavior that warranted the punishment was repeated or severe. They may also react with scolding or further punishment. That said, our model predicts positive responses in xxxxx cases, suggesting that children are successful in negotiating better outcomes at least some of the time. - costly apology model?

@@

@@

Monthly income was also not found to be a significant predictor of signaling behavior. Many models designed to explain signals of need (e.g., the begging of young birds or withdrawal in humans) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000]. However, it is possible that this can be true and household level resources have little influence on child signaling frequencies between families. As signaling behavior is costly to child, caretakers, and siblings, children may gain direct and indirect fitness benefits from limiting their signaling behavior in times of resource stress (i.e., when the chances of gaining investment or the size of investment are small). Similarly, children in high resource families may also be motivated to signal at high frequencies due to the potential for larger pieces of investment. We know of no empirical explorations of this in humans (check more and confirm). However, multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]]. ~~Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.~~ Compared to overall effort in eliciting need, resource availability may have a larger impact on how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].

@@

Our study was also predominantly cross-sectional, which limis our ability to make causal inferences . A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average. In other words, higher levels of signaling may stem from greater levels of need while simultaneously eliciting more investment from others. This hypothesized causal, bi-directional relationship makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age). Longitudinal studies can help address this. However, our small sample size of participants with signaling data and anthropometric data at two time periods prevents this for this study.

@@

## Discussion

### Running away

### Perceived immune function
